Age-related changes in plant growth and traits

### 2016-10-25 scripts/nutrient.Rmd
working up nutrient data

nutrient <- read.csv("~/Reproductive_Allocation_Kuringgai/data/nutrient_data.csv")
soil <- read.csv("~/Reproductive_Allocation_Kuringgai/data/nutrient_soil.csv")
accessoryList <- read.csv("~/Reproductive_Allocation_Kuringgai/data/accessoryList.csv")

@@

pch.tissue <- function(x=NULL){
  ret <- c(wood_from_tip=16, wood_from_middle=10,wood_from_base=7,shed_sapwood=1,bark=15,shed_bark=0,leaves=17,senescent_leaves=2,seed=8,reproductive=4)
  names(ret) <- c("wood_from_tip","wood_from_middle","wood_from_base","shed_sapwood","bark","shed_bark","leaves","senescent_leaves","seed","reproductive")
  if(!is.null(x)) {
    ret <- ret[as.character(x)]
  }
  ret
}

@@

nutrient <- nutrient %>%
  mutate(
    NP = N/P
  )

#soils data
```{r}
pdf("figs2/Nutrients/soil.pdf", height=6, width=6)
plot
plot(N~P,soil,col=col.age(age),pch=16,cex=replicate)
dev.off()
```

```{r}
nutrient_spp_age <- subset(nutrient,live_dead=="live"|live_dead=="dead")
mean_values <- function(nutrient_spp_age) {
  select(nutrient_spp_age,species,tissue,age,live_dead,K,P,N) %>%
  group_by(species,tissue,age,live_dead) %>%
  summarise_each(funs(mean),N,P,K)
}

@@

```{r}
nutrient_spp <- subset(nutrient,live_dead=="live"|live_dead=="dead")
mean_values <- function(nutrient_spp) {
  select(nutrient_spp_age,species,tissue,live_dead,K,P,N) %>%
  group_by(species,tissue,live_dead) %>%
  summarise_each(funs(mean),N,P,K)
}

@@

convert RA to nutrient values
```{r}
temp <- subset(nutrient_sum_spp,tissue=="leaves")
temp <- select(temp,species,N_dead,N_live,P_dead,P_live)
names(temp) <- c("species","N_dead_leaf","N_live_leaf","P_dead_leaf","P_live_leaf")

SummaryInd2 <- plyr::join(SummaryInd,temp, by="species")
SummaryInd2 <- SummaryInd2 %>% mutate(
  leaf_N_lost = leaf_shed*N_dead_leaf,
  leaf_P_lost = leaf_shed*P_dead_leaf,
  leaf_N_new = leaf_inv_gross*N_live_leaf,
  leaf_P_new = leaf_inv_gross*P_live_leaf,
  leaf_N_net = leaf_N_new - leaf_N_lost,
  leaf_P_net = leaf_P_new - leaf_P_lost
  )
```

Merging in reproduction data
``` {r}
InvestIndPart <- merge(InvestmentByPart,select(accessoryList,-sort), by=c("species","part"))

@@

SummaryInvest <- 
  InvestIndPart %>%
  group_by(parts_fewer_yet,individual) %>%
  summarise_each(funs(sum),weight)

SummaryInvest$species <- SummaryInvest$individual
SummaryInvest$species <- substr(SummaryInvest$species,1,nchar(SummaryInvest$species)-4)
names(SummaryInvest) <- c("tissue","individual","weight","species")

nutrient_subset <- select(nutrient_spp,species,tissue,N,P)

SummaryInvest <- merge(SummaryInvest,nutrient_subset, by=c("species","tissue"),all.x=TRUE)
names(SummaryInvest)

SummaryInvest$tissue_N <- SummaryInvest$weight*SummaryInvest$N
SummaryInvest$tissue_P <- SummaryInvest$weight*SummaryInvest$P

nutrientRepro <- 
  SummaryInvest %>%
  group_by(individual) %>%
  summarise_each(funs(sum),tissue_N,tissue_P)

SummaryInd2 <- plyr::join(SummaryInd2,nutrientRepro, by="individual")

SummaryInd2$RA_N <- SummaryInd2$tissue_N / (SummaryInd2$tissue_N + SummaryInd2$leaf_N_net)
SummaryInd2$RA_P <- SummaryInd2$tissue_P / (SummaryInd2$tissue_P + SummaryInd2$leaf_P_net)

  i <- which(is.infinite(SummaryInd2$RA_N))
  for(v in c("RA_N")) {
    SummaryInd2[[v]][i] <- NA
  }

```

xvar <- "age"
yvar3 <- "RA_N"
yvar2 <- "RA_P"
yvar <- "RA_leaf_area"

@@

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="x",ylab=yvar,cex=1,col=col.age(data[[spp]]$age),xlim=c(1,35))
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar,log="x",ylab=yvar2,cex=1,col=col.age(data[[spp]]$age),xlim=c(1,35))
  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar,log="x",xlab=xvar,ylab=yvar3,cex=1,col=col.age(data[[spp]]$age),xlim=c(1,35))
  mtext(labels.spp.full(spp),3, outer=TRUE,font=3)
  mtext("age (years)",outer=TRUE,1,line=1.5)
}

  if(spp=="BOLE"|spp=="COER"|spp=="GRSP"|spp=="HEPU") {
    rect(14,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  if(spp=="PILI") {
    rect(10,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }

plot(RA_N~age,subset(SummaryInd2,species=="BAER"),col=col.age(age),pch=16)

plot(leaf_N_lost ~ leaf_N_new,SummaryInd2,pch=16,col=col.spp(species),log="xy")
plot(leaf_N_lost ~ leaf_N_new,SummaryInd2,pch=16,col=col.age(age),log="xy")

abline(0,1)
plot(leaf_P_net ~ leaf_N_net,SummaryInd2,pch=16,col=col.spp(species),log="xy")

View(SummaryInd2)
#simple N vs P plot
```{r}
pdf("figs2/Nutrients/NvsP.pdf", height=11, width=8.5)
plot
par(mfrow=c(2,1), cex=1, omi=c(.1,.1,.05,.05), mai=c(.8,.8,.05,.05))
plot(P~N,data=subset(nutrient),col=col.spp(species),pch=pch.tissue(tissue_simple),log="xy")
legend("topleft",legend=c("sapwood","wood from mid","wood from base","shed sapwood","bark","shed bark","leaves","shed leaves", "seed", "other reproductive"),pch=c(16,10,7,1,15,0,17,2,8,4),bty="n")
legend("bottomright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

@@

```{r}
pdf("figs2/Nutrients/age_by_species_NP.pdf", height=7.5, width=11)
plot
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.05,.05), mai=c(.8,.8,.05,.05)) 
for (i in c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")) {
data <- subset(nutrient,species==i)
plot(P~N,data,col=col.age(age),pch=pch.tissue(tissue_simple),log="xy")
text(1,200,i)
data2 <- subset(data,tissue_simple=="reproductive"|tissue_simple=="seed")
text(P~N,data2,label=data2$tissue_complete,col="red",cex=0.7)
legend("bottomright",legend=labels.age(),col=col.age(),pch=16, cex=.8,bty="n")
legend("topleft",legend=c("sapwood","wood from mid","wood from base","shed sapwood","bark","shed bark","leaves","shed leaves", "seed", "other reproductive"),pch=c(16,10,7,1,15,0,17,2,8,4),bty="n")
}
dev.off()
```

#shed leaf N per year

>>>>

#starting commands
```{r}
nutrient <- read.csv("~/Reproductive_Allocation_Kuringgai/data/nutrient_data.csv")
soil <- read.csv("~/Reproductive_Allocation_Kuringgai/data/nutrient_soil.csv")
accessoryList <- read.csv("~/Reproductive_Allocation_Kuringgai/data/accessoryList.csv")

@@

pch.tissue <- function(x=NULL){
  ret <- c(wood_from_tip=16, wood_from_middle=10,wood_from_base=7,shed_sapwood=1,bark=15,shed_bark=0,leaves=17,senescent_leaves=2,seed=8,reproductive=4)

  names(ret) <- c("wood_from_tip","wood_from_middle","wood_from_base","shed_sapwood","bark","shed_bark","leaves","senescent_leaves","seed","reproductive")
  if(!is.null(x)) {
    ret <- ret[as.character(x)]
  }
  ret
}

@@

plot_abline <- function(add_lines) {
  abline(log(1),1)
  abline(log10(.1),1,col="light gray")
  abline(log10(10),1,col="light gray")
  abline(log10(.5),1,col="gray")
  abline(log10(2),1,,col="gray")
  abline(log10(.01),1,col="light blue")
  abline(log10(100),1,col="light blue")
}
```

#SUMMARIZING AND MERGING NP DATA INTO SUMMARYIND

#summarize nutrient data by species, age
```{r}
nutrient_spp_age <- subset(nutrient,live_dead=="live"|live_dead=="dead")
mean_values <- function(nutrient_spp_age) {
  select(nutrient_spp_age,species,tissue,age,live_dead,K,P,N) %>%
  group_by(species,tissue,age,live_dead) %>%
  summarise_each(funs(mean),N,P,K)
}

@@

#summarize nutrient data by species
```{r}
nutrient_spp <- subset(nutrient,live_dead=="live"|live_dead=="dead")
mean_values <- function(nutrient_spp) {
  select(nutrient_spp_age,species,tissue,live_dead,K,P,N) %>%
  group_by(species,tissue,live_dead) %>%
  summarise_each(funs(mean),N,P,K)
}

@@

#merge dataframe with species means for N,P,K content with SummaryInd
```{r}
temp <- subset(nutrient_sum_spp,tissue=="leaves")
temp <- select(temp,species,N_dead,N_live,P_dead,P_live,NUE,PUE)
names(temp) <- c("species","dead_leaf_N","live_leaf_N","dead_leaf_P","live_leaf_P","NUE","PUE")

SummaryInd2 <- plyr::join(SummaryInd,temp, by="species")
```

#merging in reproduction data
``` {r}
InvestIndPart <- merge(InvestmentByPart,select(accessoryList,-sort), by=c("species","part"))

@@

SummaryInvest <- 
  InvestIndPart %>%
  group_by(parts_to_match,individual) %>%
  summarise_each(funs(sum),weight)

SummaryInvest$species <- SummaryInvest$individual
SummaryInvest$species <- substr(SummaryInvest$species,1,nchar(SummaryInvest$species)-4)
names(SummaryInvest) <- c("parts_to_match","individual","weight","species")

nutrient_subset <- select(nutrient_spp,species,parts_to_match,N,P)

SummaryInvest <- merge(SummaryInvest,nutrient_subset, by=c("species","parts_to_match"),all.x=TRUE)
names(SummaryInvest)

SummaryInvest$repro_N <- SummaryInvest$weight*SummaryInvest$N
SummaryInvest$repro_P <- SummaryInvest$weight*SummaryInvest$P

nutrientRepro <- 
  SummaryInvest %>%
  group_by(individual) %>%
  summarise_each(funs(sum),repro_N,repro_P)

SummaryInd2 <- plyr::join(SummaryInd2,nutrientRepro, by="individual")
```

#determine leaf N,P in various tissue pools
```{r}
SummaryInd2 <- SummaryInd2 %>% mutate(
  leaf_N_lost = leaf_shed*dead_leaf_N,
  leaf_P_lost = leaf_shed*dead_leaf_P,
  leaf_N_resorbed = leaf_shed*(live_leaf_N-dead_leaf_N),
  leaf_P_resorbed = leaf_shed*(live_leaf_P-dead_leaf_P),
  leaf_N_new = leaf_inv_gross*live_leaf_N,
  leaf_P_new = leaf_inv_gross*live_leaf_P,
  leaf_N_net = leaf_N_new - leaf_N_resorbed,
  leaf_P_net = leaf_P_new - leaf_P_resorbed,
  leaf_N_for_RA = growth_leaf*live_leaf_N,
  leaf_P_for_RA = growth_leaf*live_leaf_P,
  NP_live_leaf = live_leaf_N/live_leaf_P,
  NP_dead_leaf = dead_leaf_N/dead_leaf_P,
  RA_N = repro_N / (repro_N + leaf_N_for_RA),
  RA_P = repro_P / (repro_P + leaf_P_for_RA),
  total_N_per_year = leaf_N_new + repro_N,
  total_P_per_year = leaf_P_new + repro_P,
  prop_recycled_N = leaf_N_resorbed / leaf_N_new,
  prop_recycled_P = leaf_P_resorbed / leaf_P_new,
  prop_required_N = leaf_N_resorbed / total_N_per_year,
  prop_required_P = leaf_P_resorbed / total_P_per_year
  )
```

#PLOTS

#soils data
```{r}
pdf("figs2/Nutrients/soil.pdf", height=6, width=6)
plot
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.6,.05), mai=c(.8,.8,.05,.05))
plot(N~P,soil,col=col.age(age),pch=16,cex=replicate)
dev.off()
```

#various plots using NP data in leaves, reproductive material
```{r}
pdf("figs2/Nutrients/plots_of_NP_use.pdf", height=7.5, width=11)
plot
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.6,.05), mai=c(.8,.8,.05,.05))
plot(leaf_N_net ~ repro_N,SummaryInd2, col=col.spp(species),pch=16,log="xy",xlim=c(1,10^7),ylim=c(1,10^7),xlab="",ylab="")
mtext("N in reproductive material (mg)",1, outer=FALSE,line=2)
mtext("new N from soil required for leaves (mg)",2, outer=FALSE,line=2)
plot_abline(add_lines)

plot(leaf_P_net ~ repro_P,SummaryInd2, col=col.spp(species),pch=16,log="xy",xlim=c(100,10^8),ylim=c(100,10^8),xlab="",ylab="")
mtext("P in reproductive material (ppm)",1, outer=FALSE,line=2)
mtext("new P from soil required for leaves (ppm)",2, outer=FALSE,line=2)
plot_abline(add_lines)

plot(total_N_per_year ~ leaf_N_resorbed,SummaryInd2, col=col.spp(species),pch=16,log="xy",xlim=c(1,10^7),ylim=c(1,10^7),xlab="",ylab="")
mtext("N resorbed from shed leaves (mg)",1, outer=FALSE,line=2)
mtext("total N used by plant for repro + all leaf growth (mg)",2, outer=FALSE,line=2)
plot_abline(add_lines)

plot(total_P_per_year ~ leaf_P_resorbed,SummaryInd2, col=col.spp(species),pch=16,log="xy",xlim=c(100,10^8),ylim=c(100,10^8),xlab="",ylab="")
mtext("P resorbed from shed leaves (ppm)",1, outer=FALSE,line=2)
mtext("total P used by plant for repro + all leaf growth (ppm)",2, outer=FALSE,line=2)
plot_abline(add_lines)

plot(prop_required_N ~ prop_required_P,SummaryInd2, col=col.age(age),pch=16,log="",xlim=c(-0.2,.9),ylim=c(-0.2,.9))
abline(0,1)
plot(prop_required_N ~ prop_required_P,SummaryInd2, col=col.spp(species),pch=16,log="",xlim=c(-0.2,.9),ylim=c(-0.2,.9))
abline(0,1)

plot(leaf_N_resorbed ~ leaf_N_net,subset(SummaryInd2,age>3), col=col.spp(species),pch=16,log="xy",xlim=c(1,10^7),ylim=c(1,10^7))
plot_abline(add_lines)

plot(leaf_P_resorbed ~ leaf_P_net,subset(SummaryInd2,age>3), col=col.spp(species),pch=16,log="xy",xlim=c(100,10^8),ylim=c(100,10^8))
plot_abline(add_lines)

plot(NUE~live_leaf_N,subset(SummaryInd2,LL_death<6), col=col.spp(species),pch=16,log="x",ylim=c(-.1,1))
plot(PUE~live_leaf_P,subset(SummaryInd2,LL_death<6), col=col.spp(species),pch=16,log="x",ylim=c(-.1,1))
dev.off()
```

#RA plots using NP data
```{r}
pdf("figs2/Nutrients/RA_using_NP.pdf", height=7.5, width=11)
plot
xvar <- "age"
yvar3 <- "RA_N"
yvar2 <- "RA_P"
yvar <- "RA_leaf_area"

@@

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="xy",ylab=yvar,cex=1,col=col.age(data[[spp]]$age),xlim=c(1,35))
    if(spp=="BOLE"|spp=="COER"|spp=="GRSP"|spp=="HEPU") {
    rect(14,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  if(spp=="PILI") {
    rect(10,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  plot_yvar_vs_xvar(data[[spp]], yvar2, xvar,log="xy",ylab=yvar2,cex=1,col=col.age(data[[spp]]$age),xlim=c(1,35))
      if(spp=="BOLE"|spp=="COER"|spp=="GRSP"|spp=="HEPU") {
    rect(14,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  if(spp=="PILI") {
    rect(10,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  plot_yvar_vs_xvar(data[[spp]], yvar3, xvar,log="xy",xlab=xvar,ylab=yvar3,cex=1,col=col.age(data[[spp]]$age),xlim=c(1,35))
      if(spp=="BOLE"|spp=="COER"|spp=="GRSP"|spp=="HEPU") {
    rect(14,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  if(spp=="PILI") {
    rect(10,10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="light grey",density=NA)
    rect(10^par("usr")[1],10^par("usr")[3],10^par("usr")[2],10^par("usr")[4],col="black",density=0)
  }
  mtext(labels.spp.full(spp),3, outer=TRUE,font=3)
  mtext("age (years)",outer=TRUE,1,line=1.5)
}
dev.off()
```

#simple N vs P plot
```{r}
pdf("figs2/Nutrients/NvsP.pdf", height=11, width=8.5)
plot
par(mfrow=c(2,1), cex=1, omi=c(.1,.1,.05,.05), mai=c(.8,.8,.05,.05))
plot(P~N,data=subset(nutrient),col=col.spp(species),pch=pch.tissue(tissue_simple),log="xy")
legend("topleft",legend=c("sapwood","wood from mid","wood from base","shed sapwood","bark","shed bark","leaves","shed leaves", "seed", "other reproductive"),pch=c(16,10,7,1,15,0,17,2,8,4),bty="n")
legend("bottomright",legend=labels.spp(),col=col.spp(),pch=16, cex=.8,bty="n")

@@

#plotting NP data for each species, colored by age
```{r}
pdf("figs2/Nutrients/age_by_species_NP.pdf", height=7.5, width=11)
plot
par(mfrow=c(1,2), cex=1, omi=c(.1,.1,.05,.05), mai=c(.8,.8,.05,.05)) 
for (i in c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")) {
data <- subset(nutrient,species==i)
plot(P~N,data,col=col.age(age),pch=pch.tissue(tissue_simple),log="xy")
text(1,200,i)
data2 <- subset(data,tissue_simple=="reproductive"|tissue_simple=="seed")
text(P~N,data2,label=data2$tissue_complete,col="red",cex=0.7)
legend("bottomright",legend=labels.age(),col=col.age(),pch=16, cex=.8,bty="n")
legend("topleft",legend=c("sapwood","wood from mid","wood from base","shed sapwood","bark","shed bark","leaves","shed leaves", "seed", "other reproductive"),pch=c(16,10,7,1,15,0,17,2,8,4),bty="n")
}
dev.off()
```
